---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
  html_document: default
---

**Principles of Data Visualization and Introduction to ggplot2**

I have provided you with data about the 5,000 fastest growing companies in the US, as compiled by Inc. magazine. lets read this in:

```{r}
inc <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module1/Data/inc5000_data.csv", header= TRUE)
```

And lets preview this data:

```{r}
head(inc)
```
```{r}
summary(inc)
```

Think a bit on what these summaries mean. Use the space below to add some more relevant non-visual exploratory information you think helps you understand this data:

Below I used the skim function from the skimr library. It gives a nice clean display of the dimensions, data types, missing values, and a quick histogram for each variable. We can see below that the only field with missing values is the Employees field, it is only missing 12 out of 5001 records so we will need to take this into account when asnwering question 2. The histograms show a uniform distribution for Rank and a a uniform distribution for growth rate, revenue and employees with a smaller range.

```{r}

library(skimr)
skim(inc)

```

## Question 1

Create a graph that shows the distribution of companies in the dataset by State (ie how many are in each state). There are a lot of States, so consider which axis you should use. This visualization is ultimately going to be consumed on a 'portrait' oriented screen (ie taller than wide), which should further guide your layout choices.



```{r}
library(ggplot2)

state_count<- inc %>% group_by(State) %>% tally()

p <- ggplot(state_count, aes(x = n, y= reorder(State, n))) +
        geom_bar(stat = "identity",width=1,fill='light blue')+ theme(aspect.ratio=16/9)+
        theme(text = element_text(size=8),
        axis.text.y = element_text(angle=0, hjust=1))  +
        geom_text(size=1.9,aes(label=n), position=position_dodge(width=.25), hjust=1)+
        theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))+
        xlab("Count of Companies")+
        ylab("State")+
        ggtitle("5,000 Fastest Growing Comapnies in the US by State")

p


```

## Quesiton 2

Lets dig in on the state with the 3rd most companies in the data set. Imagine you work for the state and are interested in how many people are employed by companies in different industries. Create a plot that shows the average and/or median employment by industry for companies in this state (only use cases with full data, use R's `complete.cases()` function.) In addition to this, your graph should show how variable the ranges are, and you should deal with outliers.

Some of the companies in the dataset for New York have some extreme outliers. I used the boxplot function to identify outliers and remove them from my population. I then used the Cleveland dot-plot code as found in section 3.10 in the r cookbook in order to show a general idea of the means and medians employees per industry. Then I plotted a boxplot by industry to show the range and median and then I plotted a point for the mean so that could be compared to the median as well.

```{r}
library(ggplot2)

complete_inc<-inc[complete.cases(inc), ]
inc_ny <- filter(complete_inc, State == 'NY')

outliers <- boxplot(inc_ny$Employees, plot=FALSE)$out
inc_ny<- inc_ny[-which(inc_ny$Employees %in% outliers),]

ny_mean<-aggregate(x = inc_ny$Employees,                
          by = list(inc_ny$Industry),             
          FUN = mean)
ny_mean$type<-'mean'

ny_med<-aggregate(x = inc_ny$Employees,                
          by = list(inc_ny$Industry),             
          FUN = median)

ny_med$type<-'median'

ny_avg<-rbind(ny_mean,ny_med)



ggplot(ny_avg, aes(x = x, y = reorder(Group.1,x))) +
  geom_segment(aes(yend = Group.1), xend = 0, colour = "grey50") +
  geom_point(size = 3, aes(colour = type)) +
  scale_colour_brewer(palette = "Set1", limits = c("mean", "median")) +
  theme_bw() +
  theme(
    panel.grid.major.y = element_blank(),   # No horizontal grid lines
    legend.position = c(1, 0.55),           # Put legend inside plot area
    legend.justification = c(1, 0.5)
  )+xlab("Average Number of Employees")+
        ylab("Industry")+
        ggtitle("Average Number of Employees by Industry in NY")

ggplot(inc_ny, aes(x = factor(Industry), y = Employees)) +
  geom_boxplot() +
  stat_summary(fun = "mean", geom = "point", shape = 23, size = 3, fill = "white")+
   theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    axis.text.x = element_text(angle = 60, hjust = 1)
  ) +
  xlab("Industry")+
        ylab("Number of Employees")+
        ggtitle("Box-Plot of the Number of Employees by Industry in NY")
```

## Question 3

Now imagine you work for an investor and want to see which industries generate the most revenue per employee. Create a chart that makes this information clear. Once again, the distribution per industry should be shown. 

I followed the same outline as question 2 where I removed outliers of the key field. In this case I used a violin plot so I could very easily see the distribution and it would be even more easily read than a box-plot. For example we can see that Logistics and Transportation has a more uniform distribution then the other industries. Based on this limited statistic it seems that Computer Hardware and Manufacturing offer the most bang for your buck per employee.

```{r}

complete_inc$rev_per_emp<-round(complete_inc$Revenue/complete_inc$Employees,2)

outliers <- boxplot(complete_inc$rev_per_emp, plot=FALSE)$out
complete_inc<- complete_inc[-which(complete_inc$rev_per_emp %in% outliers),]


ggplot(complete_inc, aes(x = factor(Industry), y = rev_per_emp)) +
  geom_violin(fill='Light Grey') +
   theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    axis.text.x = element_text(angle = 60, hjust = 1)
  ) +   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))+
  xlab("Industry")+
        ylab("Revenue per Employee in US $")+
        ggtitle("Revenue per Employee in US $ by Industry")
```



```
In addition to the reading material for the course I used the resources below

Links:

https://stackoverflow.com/questions/16961921/plot-data-in-descending-order-as-appears-in-data-frame
https://stackoverflow.com/questions/12018499/how-to-put-labels-over-geom-bar-for-each-bar-in-r-with-ggplot2
https://felixfan.github.io/ggplot2-remove-grid-background-margin/
https://stackoverflow.com/questions/18252827/increasing-area-around-plot-area-in-ggplot2

https://www.r-bloggers.com/2020/01/how-to-remove-outliers-in-r/
https://statisticsglobe.com/mean-by-group-in-r
